Peaks Over Threshold for Bursty Time Series

### 2018-04-16 article_springer/article_springer.Rmd
proof with details on subordination

@@

*Proof:*
Let the scaling functions $a(c)$ and $d(c)$ be as in Section 
\ref{magnitudes-and-extreme-value-distributions}, and assume WLOG 
$\ell = a(n) \ell^* + d(n)$ for some arbitrary but fixed $\ell^*$, out of the support of $A$. 
We first note that due to \eqref{eq:GEV}
\begin{equation} 
F_J (a(c) \ell^* + d(c))^c \stackrel{d}{\to} F_A(\ell^*), \text{ as } c \to \infty.
\end{equation}
\noindent Taking the logarithm on each side and using the relationship  $z-1 \sim \log(z)$ as $z \to 1$, we get
\begin{equation}
c (1-F_J(a(c) \ell^*+d(c))) \stackrel{d}{\to} - \log F_A(\ell^*).
\end{equation}
\noindent Hence we have 
\begin{equation}
c \cdot p \stackrel{d}{\to} - \log F_A(\ell^*) \label{eq:np_conv}
\end{equation}
\noindent as $\ell \uparrow x_R$.
Now we get
\begin{align}
\begin{split}
\mathbf P \left[T(\ell)/b(1/p) > t \right]
&= \mathbf P[V(b(1/p) t) \leq \ell] \label{eq:proofTheo}
= \mathbf P[V(b(1/p)) \le a(c) \ell^* + d(c)]\\
&= \mathbf P\left(\frac{V(b(1/p)/b(c) \cdot b(c) \cdot t)-d(c)}{a(c)} \le \ell^*\right) \\
&\stackrel{d}{\to} \mathbf P(A(E(-\log F_A(l))^{-1/\beta} t) \leq \ell^*),
\end{split}
\end{align}
where the convergence follows due to Theorem MS1 (with $c$ replaced by $b(c)$) and the fact that
\begin{equation}
\frac{b(1/p)}{b(c)} 
= \frac{b(1/cp \cdot c)}{b(c)} 
\sim (cp)^{-1/\beta}
\to \left( -\log F_A(l)\right)^{-1/\beta} 
\text{ as } c \to \infty.\label{eq:proofTheo2}
\end{equation}
In \eqref{eq:proofTheo2} we used \eqref{eq:np_conv} and 
$b \in RV(1/\beta)$, which means that 
$b(\lambda c) / b(c) \rightarrow \lambda^{1/\beta}$ as 
$c \rightarrow \infty$ uniformly in $\lambda$ on each compact subset of 
$(0,\infty)$. 
The limit in \eqref{eq:proofTheo} is equal to
$$\mathbf P \left( \left( -\log(F_A(\ell^{*})) \right)^ {1/\beta} B(\ell^*) > t\right).$$ 
Due to Theorem MS2 and \eqref{eq:hitting-ML}, we know that 
$(-\log F_A (\ell^*))^{1/\beta} B(\ell^*)$ is $ML(\beta,1)$ distributed, 
and hence the theorem follows.

>>>>

_Proof of Theorem:_
We interpret the threshold crossing time as the hitting time of the 
underlying CTRM (Continuous Time Random Maxima) or "max-renewal process", 
and utilize a result by @MeerschaertStoev08. 
The running maximum process is defined as 
$$
M_J(t) := J_1 \vee \ldots \vee J_{\lfloor t \rfloor},
$$
and as we assume the $J_k$ have a continuous distribution, there exist 
norming functions $a(c)$ and $d(c)$ such that 
$$
\frac{M_J(t) - d(t)}{a(t)} \stackrel{d}{\longrightarrow} A, \quad t \to \infty
$$
where $A$ is a random variable with a generalized extreme value distribution. 
The underlying CTRM is then defined via 
$$
X(t) = M_J(N(t)), \quad t \ge 0
$$
where $N(t)$ is the renewal process associated with the waiting times 
$W_k$: 
$$
N(t) = \max\{n: W_1 + \ldots + W_n \le t\}.
$$
Now a key observation is that
$$
T(\ell) = \inf\{t: V(t) > \ell\}, 
$$
and that 
$$
T(\ell) > t \quad \text{ if and only if } \quad V(t) \le \ell.
$$
By [Theorem 3.1, @MeerschaertStoev08], we have the stochastic process 
convergence 
$$
\frac{V(ct) - d(\tilde b(c))}{a(\tilde b(c))} \stackrel{d}{\longrightarrow} 
A(E(t)), \quad t > 0.
$$
where:

* $A(t)$ is the extremal process such that $A(1) \stackrel{d}{=} A$, 
* $E(t) = \inf\{u: D(u) > t\}$ is an inverse subordinator, 
  associated with the Levy process $D(t)$ that satisfies 
  $D(1) \stackrel{d}{=} D$,
* $\tilde b(c)$ is a regularly varying norming function which is
  _inverse_ to $b(c)$: $b(\tilde b(c)) \sim c \sim \tilde b(b(c))$.

Now 
\begin{align}
\begin{split}
\mathbf P \left[T(\ell)/b(1/p) > t \right]
&= \mathbf P[V(b(1/p) t) \leq \ell] \label{eq:proofTheo}
= \mathbf P[V(b(1/p)) \le a(c) \ell^* + d(c)]\\
&= \mathbf P\left(\frac{V(b(1/p)/b(c) \cdot b(c) \cdot t)-d(c)}{a(c)} \le \ell^*\right) \\
&\stackrel{d}{\to} \mathbf P(A(E(-\log F_A(l))^{-1/\beta} t) \leq \ell^*),
\end{split}
\end{align}
where the convergence follows due to Theorem MS1 (with $c$ replaced by $b(c)$) and the fact that
\begin{equation}
\frac{b(1/p)}{b(c)} 
= \frac{b(1/cp \cdot c)}{b(c)} 
\sim (cp)^{-1/\beta}
\to \left( -\log F_A(l)\right)^{-1/\beta} 
\text{ as } c \to \infty.\label{eq:proofTheo2}
\end{equation}
In \eqref{eq:proofTheo2} we used \eqref{eq:np_conv} and 
$b \in RV(1/\beta)$, which means that 
$b(\lambda c) / b(c) \rightarrow \lambda^{1/\beta}$ as 
$c \rightarrow \infty$ uniformly in $\lambda$ on each compact subset of 
$(0,\infty)$. 
The limit in \eqref{eq:proofTheo} is equal to
$$\mathbf P \left( \left( -\log(F_A(\ell^{*})) \right)^ {1/\beta} B(\ell^*) > t\right).$$ 
Due to Theorem MS2 and \eqref{eq:hitting-ML}, we know that 
$(-\log F_A (\ell^*))^{1/\beta} B(\ell^*)$ is $ML(\beta,1)$ distributed, 
and hence the theorem follows.

# Statistics of the Mittag-Leffler distribution

@@

*Proof:*
Let the scaling functions $a(c)$ and $d(c)$ be as in Section 
\ref{magnitudes-and-extreme-value-distributions}, and assume WLOG 
$\ell = a(n) \ell^* + d(n)$ for some arbitrary but fixed $\ell^*$, out of the support of $A$. 
We first note that due to \eqref{eq:GEV}
\begin{equation} 
F_J (a(c) \ell^* + d(c))^c \stackrel{d}{\to} F_A(\ell^*), \text{ as } c \to \infty.
\end{equation}
\noindent Taking the logarithm on each side and using the relationship  $z-1 \sim \log(z)$ as $z \to 1$, we get
\begin{equation}
c (1-F_J(a(c) \ell^*+d(c))) \stackrel{d}{\to} - \log F_A(\ell^*).
\end{equation}
\noindent Hence we have 
\begin{equation}
c \cdot p \stackrel{d}{\to} - \log F_A(\ell^*) \label{eq:np_conv}
\end{equation}
\noindent as $\ell \uparrow x_R$.
Now we get
